# Publication

Projects generate a variety of types of publications. These may include

- Pre-analysis plans,
- Final technical reports, 
- Public versions of GitHub repositories, 
- Public versions of data, and
- Peer-reviewed academic journal articles. 

## Authorship in Pre-Analysis Plans

For pre-analysis plans, we ascribe to the three standards of _substantial contribution_, _approval_, and _accountability_ that _Nature_ lays out [here](https://www.nature.com/nature/editorial-policies/authorship).


## Authorship in Academic Articles

For academic papers, we ascribe to the three standards of _substantial contribution_, _approval_, and _accountability_ that _Nature_ lays out [here](https://www.nature.com/nature/editorial-policies/authorship).

## Publishing Data

We strive to post useful replication data to the extent possible in each project. Some examples of our work doing so include projects on the 911 [nurse triage line](https://github.com/thelabdc/FEMS-911NurseTriageLine-Public) and 457b [enrollment](https://github.com/thelabdc/DCHR-457b-Public).

The federal OPRE has some helpful advice for doing so [here](https://www.acf.hhs.gov/opre/report/guide-publishing-research-data-secondary-analysis).

## Presenting Results

### Estimates and uncertainties

When we want to communicate group estimates along with treatment effects, we display them with barplots or scatterplots. Where we are interested in inference about a treatment effect, we display the control group and treatment group means, with the uncertainty around the treatment group mean representing the uncertainty around the treatment effect. For example, 

```{r display_te_uncertainty, echo=FALSE, message=FALSE}
library(estimatr)
library(tidyverse)

set.seed(739696075)

n_te_display <- 100


df_te_display <- tibble(tr = sample(0:1, n_te_display, replace = TRUE),
                        y = 5 + tr + rnorm(n_te_display))

lm_te_display <- lm_robust(y ~ tr, data = df_te_display) 

lm_te_display_tidy <- lm_te_display |> tidy()

mean_co <- lm_te_display_tidy |> 
  filter(term == "(Intercept)") |> 
  pull(estimate)

te <- lm_te_display_tidy |> 
  filter(term == "tr") |> 
  pull(estimate)

mean_tr <- mean_co + te

te_se <- lm_te_display_tidy |> 
  filter(term == "tr") |> 
  pull(std.error)

df_means <- tibble(Group = c("Control", "Treatment"),
                   Mean = c(mean_co, mean_tr),
                   se = c(NA, te_se))

ggplot(df_means, aes(x = Group, y = Mean, fill = Group)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(data = df_means[df_means$Group == "Treatment", ],
                aes(ymin = Mean - 1.96 * se, 
                    ymax = Mean + 1.96 * se),
                width = 0.3) +
  theme_minimal()
```

When we want to communicate treatment effects without both group means being explicit, we do so with a coefficient plot. For example, with the intercept (control group mean),

```{r display_coefplot, echo=FALSE, message=FALSE}
library(modelsummary)
library(patchwork)
mp1 <- modelplot(lm_te_display) +
  geom_vline(xintercept = 0) + 
  lims(x = c(0, 6))
mp1
```

or without the intercept,


```{r display_coefplot_noint, echo=FALSE, message=FALSE}
mp2 <- modelplot(lm_te_display, coef_omit = 'Intercept') +
  geom_vline(xintercept = 0) +
  lims(x = c(0, 3))
mp2
```

