# (APPENDIX) Appendix {-} 

# R Basics

## Get R and RStudio

First download and install R from https://cran.r-project.org.

Then, download and install RStudio (Desktop) from https://posit.co/products/open-source/rstudio/.

(You should have permission for both of these. If not, you may need to contact
EOM IT.)

## Read in data files

We can read data from many formats and many places, including from a local file
or from the web. Many R packages contain their own data.

The code below requires the following packages to be loaded and attached:

```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(here)
library(lubridate)
library(magrittr)
library(readr)
```

### `.csv`

Data from the web (from the résumé audit experiment from @bermul04):

```{r message=FALSE}
df_resume <- read_csv("https://raw.githubusercontent.com/kosukeimai/qss/master/CAUSALITY/resume.csv")
```

Then we can examine the dimensions of the new object named `df_resume`:

```{r}
dim(df_resume)
```

So, `df_resume` has `r nrow(df_resume)` rows and `r ncol(df_resume)` columns. We
can see the first few rows with

```{r}
head(df_resume)
```


For data from a local file, if data `mydata.csv` are in subdirectory `/data/`,  

```{r eval=FALSE}
my_data_path <- here("data", "mydata.csv")

df <- read_csv(my_data_path)
```

See
[here](https://thelab.dc.gov/LAB-SOP-experiments/appendix-b-managing-code.html#working-directory-and-relative-paths)
for more detail on how to use `here()` to create paths.

### `.xlsx`

Use package `readxl` and function `read_excel()`. To read the second sheet from
a multi-sheet spreadsheet, add an argument like `read_excel(..., sheet = 2)`.


### Box

To read a file directly from box, we use the `boxr` package. To do so, 

- Have a member of the data science team (currently, Nathan) add you as Box RStudio App collaborator
- Go to dcgov.box.com. At the bottom left, click "Dev Console"
- Click on The Lab @ DC
- Click the Configuration tab

Here you should see a `client_id` and `client_secret` that you can copy.

Then, run at the R Console

```{r eval=FALSE}
library(boxr)

box_auth(client_id = "<INSERT ID STRING HERE>",
         client_secret = "<INSERT SECRET STRING HERE>")
```

replacing the strings with the information from the Configuration tab.

If R prompts you to update the Renviron file to store this information for next time, you may do so.

Then, your `.R` file that reads the data should include

```{r eval=FALSE}
library(boxr)

box_auth() 
# (arguments can be empty, if credentials stored in Renviron)

df <- box_read("791112121820")
# (where 791112121820 is the Box ID for a .csv to read)
```

For more information, see the `boxr` [vignette](https://r-box.github.io/boxr/articles/boxr.html).

## Pipes

Pipe operators make it easier for humans to follow a code chunk, and they allow
us to string together a sequence of operations into a "pipeline". The native R pipe is `|>`; the pipe from package `magrittr` is also very common (`%>%`).

The pipe simply takes the argument on its left and passes it as the first
argument of the function to its right. So, if I have function `f()` to which I
will pass arguments `x` and `y` (in that order), I can write

```{r eval=FALSE}
f(x, y)
```

Using a pipe, I can equivalently write either

```{r eval=FALSE}
x |> f(y)
```

or

```{r eval=FALSE}
x %>% f(y)
```


## Rename variables

Variable names should adhere to good  [practices](https://thelab.dc.gov/LAB-SOP-experiments/managing-code.html#best-practices). 

The `rename()` function in `dplyr` can rename several variables at once:

```{r eval=FALSE}
df <- df |> rename(new_var_name_1 = `Bad, old, long, spacey name`,
                    new_var_name_2 = "inconsistentsquishedvarname",
                    new_var_name_3 = starts_with("Unique string old name starts with"))
```

To automatically clean up all the names in a dataframe, the `{janitor}` package can help:

```{r eval=FALSE}
df <- df |> clean_names()
```



## Create a New Variable

Suppose I want to create a new variable and attach it to the data frame (i.e.,
to update the data frame to include the new variable). Mechanically, in the
tidyverse I overwrite the old data frame with the new one. Sometimes we want to
use a new name for the augmented data frame, but not when we create each new
variable.

### Create a "wave ID"

A "wave ID" is a variable that is constant for every row in the data set.
Suppose that the resume data were all from "wave 3" of a larger study. To
identify them as such,

```{r}
df_resume <- df_resume |> mutate(wave = 3)
```

### Create a primary key

A "primary key" is a variable that uniquely identifies each row in the data set.
To create a primary key or "ID variable",

```{r}
df_resume <- df_resume |> mutate(id_var = row_number())
```


### Transform an existing variable

Below, we use the `mutate()` function with the character variable `sex`, which
has levels `female` and `male`, to create a logical `TRUE`/`FALSE` variable
indicating whether the résumé had a putatively female name at the top:

```{r}
df_resume <- df_resume |> mutate(isFemale = sex == "female")
```

Now, `df_resume` has `r ncol(df_resume)` columns:

```{r}
head(df_resume)
```

Note that `mutate()` needs a dataframe as its first argument. Above, we pass
`df_resume` to `mutate()` using the pipe.

### Create a sum from a subset of variables

Below, we select a subset of columns from `df_resume`, take the sum across them, and add the sum as a new variable. Here, we locate the new variable right after those columns that it summed.

```{r}
df_resume |> 
  mutate(new_sum = rowSums(select(df_resume, call:id_var))) |> 
  relocate(new_sum, .after = id_var)
```



## Recode a variable's values

Instead of creating a new variable as above, we can use `mutate()` and `case_when()` to recode a variable. Suppose we wanted to code race:

```{r}
df_resume |> mutate(
  race = case_when(
    race == "white" ~ "race_wh",       # recode "white" to "race_wh"
    race == "black" ~ "race_bl",       # recode "black" to "race_bl"
    race == "missing" ~ NA_character_, # recode "missing" to (character) NA
    TRUE ~ race                        # recode any other value to original value of race
  )
)
```

There is also `recode()`, but its lifecycle status is "questioning":

```{r}
df_resume |> mutate(
  race = recode(race,
    "white" = 0,
    "black" = 1
    )
)
```


## Treat dates as dates

When we have dates stored as character strings, we transform them to dates. First, simulate some date data:

```{r}
# Simulate date data:
set.seed(117894985)

df_resume <- df_resume |> 
  mutate( 
    dob = paste0(sample(1:12, n(), replace = TRUE), 
                 "/", 
                 sample(1:28, n(), replace = TRUE),
                 "/",
                 sample(1950:2000, n(), replace = TRUE))
)

head(df_resume, 3)
```

The tidyverse includes the `lubridate` package. If you know the format of the character strings, there are many shortcut parsers like `mdy()`. Below, we recode character dates to be of date type:

```{r}
df_resume |> mutate(
  dob_date = mdy(dob)
)
```

If we aren't sure of the format, or there are several, we use `parse_date_time()` and specify several "orders" of date format. `parse()` functions do their best to use the information in "orders". Below, `"dmy"` alone would recode only the ambiguous dates, assuming they were in day-month-year order. Adding `"%m/%d/%Y"` instructs `parse()` to look for `month/day/year` formats. Finding many that are unambiguously that format, it assumes that format for the new variable. 

```{r}
df_resume |> mutate(
  dob_date = as_date(    
    parse_date_time(dob, 
                    orders = c("dmy", "%m/%d/%Y"))
  )
)
```

```{r echo=FALSE}
df_resume %<>% mutate(
  dob_date = mdy(dob)
)
```


### Make an age variable 

To make an age variable from the year component, 

```{r}
df_resume |> mutate(
  age = 2023 - year(dob_date)
)
```

For more precision, the participants' ages as of 2 May 2023, 

```{r}
df_resume |> mutate(
  age = as.numeric(
    interval(dob_date, "2023-05-02"), 
    "years")
)
```

(To get participants' ages today, replace `"2023-05-02"` above with `today()`.)

### Make a month-count variable

We may have dates accurate to the day, but we want a count of which month of
participation it is for each participant. Assume that each `firstname` in the
`resume` data represents a participant, and each `dob_date` represents an event
for that participant that we care about. We want to produce a variable that is 0 for the first month of participation (say, February), and 2 for the month two months later (April), regardless of whether the month between (March) appears in the data, and even if participants start or finish participation at very different times.^[If, instead, we want a variable with 1 for the first date for that participant, 2 for the second date, etc., _regardless of how far between the two dates are_, we can use `dplyr::dense_rank()`.]

Below, we group by `firstname`, then round all months to the first date of the month, then count those months (within `firstname`). We show the first three rows for each participant.

```{r}
df_resume |> 
  # Ensure ranking _within_ firstname:
  group_by(firstname) |> 
  # Round all months to first date of the month, then count:
  mutate(dob_month = floor_date(dob_date, unit = "month"), 
         dob_month_count = (min(dob_month) %--% dob_month) %/% months(1)
         ) |>
  # For presentation, select, arrange, and slice:
  select(firstname, dob, dob_date, dob_month, dob_month_count) |> 
  arrange(firstname, dob_month_count) |> 
  slice(1:3)
```

If we prefer the first month to be labeled `1` rather than `0`, we just replace
`%/% months(1)` with `%/% months(1) + 1`.

## Create a completely randomised indicator {#sec-create-treatment}

This uses `mutate()`, along with the `sample()` function to create a random
vector of `Treatment`'s and `Control`'s and attach it to `df_resume`:

```{r}
set.seed(317334706)

df_resume <- df_resume |> mutate(
  my_new_treatment_assg = sample(c("Treatment", "Control"), 
                                 size = n(), 
                                 replace = TRUE))
```

```{r}
head(df_resume |> select(-isFemale, -dob))
```

See also the `{randomizr}` package for a variety of helpers.

## Plot a variable

Use `ggplot()` to plot the distribution of calls back that these résumés
receive. Calls are fairly rare.

```{r barplot_resume}
df_resume |> ggplot(aes(x = call)) + geom_bar()
```

Look at the distribution by the putative race on the resume:

```{r barplot_resume_2}
df_resume |> 
  ggplot(aes(x = call, fill = race)) + 
  geom_bar(position = "dodge2") 
```

Since there are exactly the same number of résumés with each race, these count
distributions are comparable.

## Calculate a data summary

### Proportions for a categorical variable

```{r}
df_resume |> count(firstname) |> mutate(proportion = n / sum(n))
```

### Proportions for binary variables

```{r}
summ_resume <- df_resume |> 
  group_by(race) |>
  summarise(call_back_rate = mean(call),         # mean of numeric 0/1
            prop_female = mean(sex == "female")) # mean of logical TRUE/FALSE

summ_resume
```

```{r echo=FALSE}
te <- summ_resume$call_back_rate[2] - summ_resume$call_back_rate[1]
bal <- summ_resume$prop_female[1] - summ_resume$prop_female[2]
```


The call-back rates differ by about `r round(te, 3)`; the proportion female, which will be balanced by the randomization on average, only differs by about `r round(bal, 3)`.

## Merge (join) dataframes

For a fuller treatment of merges ("joins"), see [here](https://r4ds.had.co.nz/relational-data.html#relational-data) in @wicgro16. 

The four main mutating join commands are `left_join()`, `right_join()`, `inner_join()`, and `full_join()`. 

* `left_join(x, y)` keeps all rows of `x`, but only rows of `y` that have matches in `x`
* `right_join(x, y)` keeps all rows of `y`, but only rows of `x` that have matches in `y`
* `inner_join(x, y)` only keeps rows that appear in both `x` and`y`
* `full_join(x, y)` keeps all rows of both `x` and `y`

For any columns that didn't appear in the original dataframe for a certain row, the joins will store `NA`.

By default, joins will match on any columns that have the same name. So, if `age` and `height` are both in `df1` and `df2`, then `full_join(df1, df2)` will consider a row "matched" if the values of both `age` and `height` are the same in the datasets. 

If column names differ, the `by = ...` argument can inform the join. E.g., if, instead, `df1` has variables `Age` and `Tallness`, we could join with 

```{r fulljoin, eval=FALSE}
df_joined <- df1 |> full_join(df2, 
                               by = c("Age" = "age",
                                      "Tallness" = "height"))
```

Columns with the same name, but different types, will not join. E.g., if `df1$this_var` is numeric, but `df2$this_var` is character, the joins will throw an error.

## Anonymize sensitive data

Sometimes we want to preserve features of sensitive data, but not the values themselves. For example, we want to match individuals across datasets, but do not want to preserve the individuals' addresses.

```{r sensitivedata, warning=FALSE, echo=FALSE}
sens <- tibble(Name = rep(c("Beth", "Esme", "Charles")),
               Address = c("10 Downing St",
                           "667 Dark Ave",
                           "10 Downing St"),
               Phone = c(123, 456, 789),
               Score = c(90, 50, 70))
```

Suppose we have the data below, where each individual has a unique name and phone number, but two share an address:

```{r warning = FALSE, echo=FALSE}
sens
```

We can anonymize the `Address` and `Phone` columns in this data set (called `sens`) using the `digest()` function in package `{digest}`:

```{r hash, warning = FALSE}
library(digest)

cols_to_mask <- c("Address", "Phone")
for(i in cols_to_mask){
  anon <- sapply(unlist(sens[, i]), digest, algo = "sha512")
  
  # Shorten anonymized column (only for presentation here):
  short_anon <- substr(anon, 1, 10)
  
  sens[, i] <- short_anon
}

sens
```

Note that addresses 1 and 3 are the same, despite being hashed.

### Salting

Our default one-way procedure is to salt, hash, and key-stretch our sensitive data. Below is an example for a single value we wish to obscure.

```{r hashsalt, warning=FALSE, message=FALSE}
library(openssl)

sens <- "1234567"

bcrypt_pbkdf(sens,
             salt = raw(849025))
```

Above, `849025` is a value obtained from `sample(1e6, 1)` that is used to create the raw salt vector. This value should change for each value we want to hash.
