--- 
title: "SOPs for The Lab @ DC"
author: "Ryan T. Moore"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
github-repo: rtm-dc/LAB-SOP-experiments
url: 'https\://rtm-dc.github.io/LAB-SOP-experiments/'
description: "Standard Operating Procedures for Experiments and Data Analysis at The Lab @ DC."
code_folding: show
---

# Preface {.unnumbered}

This provides statistical guidance for randomized experiments, observational causal inference, and data analysis conducted at The Lab \@ DC.

The appendices provide examples of some fundamental operations in R and some standards for managing code and workflow. These examples intend to serve as templates that you can copy, alter, and use in your project work.

Please also see our developing data science standard operating [procedures](https://github.com/thelabdc/labDsSops) and, perhaps, our older [wiki](https://github.com/thelabdc/LAB-Wiki/wiki).

## Colophon {.unnumbered}

This version of the book was built with

```{r message = FALSE, warning = FALSE, echo = FALSE}
library(dplyr)

all_pkgs <- renv::dependencies(progress = FALSE) |> 
 pull(Package) 

session <- devtools::session_info(pkgs = all_pkgs)

session$platform
```

Along with these packages:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session$packages |> 
  select(package, loadedversion, date, source) |> 
  DT::datatable(rownames = FALSE,
                class = 'cell-border stripe',
                filter = list(position = 'top'),
                options = list(pageLength = nrow(session$packages), 
                           autoWidth = TRUE,
                           bInfo = FALSE,
                           paging = FALSE))
```

The code for building this colophon comes from Jenny Bryan's book [here](https://github.com/rstudio-education/stat545/blob/master/index.Rmd).
